---
title: "Example Breaking News Article Using Inline R Code"
editor: visual
---

# Stock Drop Leaves Wallstreet Shocked

```{r setup, include=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidyquant)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

```

```{r, include=FALSE, message=FALSE}

ticker <- "DIS"

stock1 = tq_get(ticker, get = "stock.prices", from = "2023-05-08")

stock2 = tq_get(ticker, get = "stock.prices", from = "2023-05-11")

mean_stock1_close = mean(stock1$close)

mean_stock2_open = mean(stock2$open)

mean_stock2_close = mean(stock2$close)

```

The Walt Disney Company suffered a major drop in stock today, leaving many experts confused. DIS, as it's known by on the stock market, was up at `r mean(stock1$close)` points at closing on Monday this week. DIS remained relatively high at opening as well today, at `r mean(stock2$open)` points. However, the stock had dropped to `r mean(stock2$close)` points by closing, leaving experts confused.

In an interview with Danny Smith, an analyst at Goldman Sachs, Smith expressed total shock over this development, "you just don't expect Disney to get hit. Every other company, fine---couldn't care less, really. But Disney? They make children's hearts sing. And for what?"

Smith's outrage, however, is not matched by other pockets of the population. Elsewhere in New York City, the Writer's Guild of America, or WGA, has been on strike due to claims of drops in pay, and other changes in the industry as streaming services, such as Netflix and Disney+, have overtaken cable television.

When told what Smith had said, Jen Fields, writer on Disney+'s third claymation remake of Beauty & the Beast, said, "how can anybody be surprised about the whole stock thing? I am about to lose my job to ChatGBT. Tell Danny Smith, if anyone deserves to lose their job to a robot, it's him"

Fields is not alone in this assessment. When prompted, many other writers hailed insult upon insult at Smith. One Twitter user commented, "his hair is so loser-core, I hope disney stock plummets to hell to join him."
