---
title: "map"
editor: visual
---

Here I will be taking you through how I am creating an interactive map displaying the median income in Massachusetts, using census data.

```{r}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tigris)
library(sf)
library(tidycensus)
library(htmltools)
library(janitor)
library(here)
library(mapview)
library(leafsync)
library(leaflet.extras2)
options(tigris_class = "sf")


```

First, I will be coding a few variables of interest. We will be specifically analyzing the median income variable.

```{r}
myvars <- c(totalpop = "B01003_001",
            medincome = "B19013_001",
            medage = "B01002_001"
)
```

Then, I will retrieve the data from the American Community Survey feature, using tidycensus in R.

```{r}
ma_counties_withgeo <- get_acs(geography = "county",
                       variables = c(myvars),
                       state = "MA",
                       output = "wide",
                       geometry = TRUE)

ma_counties_withgeo
```

Now I'll just clean up the data a bit and get it prepped for the map.

```{r}
ma_counties_withgeo <- ma_counties_withgeo %>%
  select(-ends_with("M"))
```

```{r}

colnames(ma_counties_withgeo) <- sub("E$", "", colnames(ma_counties_withgeo))

```

And behold! Using mapview, below I can display an interactive map displaying the median income for each county in the state of Massachusetts.

```{r}
mapview(ma_counties_withgeo, zcol = "medincome", 
         col.regions = RColorBrewer::brewer.pal(9, "Greens"), 
         alpha.regions = 1)
```
